\input{preamble}

% OK, start here.
%
\begin{document}

\title{Introduction}


\maketitle

\phantomsection
\label{section-phantom}


\tableofcontents


\section{Overview}
\label{section-overview}

\noindent
[This is a test, again!]

\cite{reference}

The Automorphic Project was inspired by the success and incredible utility of
the Stacks Project, and necessitated by the lack of a comprehensive 
introduction to automorphic forms and the Langlands program. While there are,
 and continue to appear, many good references on many aspects of the subject, 
we felt that it can be much better presented, in all its vastness, by an 
evolving and collaborative effort that today's technology enables.

While we will eventually aim at making the Automorphic Project as inclusive as
 possible, our goal is not to reproduce the existing literature. Rather, and
 taking into account that the literature in the field is usually considered 
very technical and difficult to follow for outsiders, our goal is to recast 
the literature in a way as conceptual as possible, and a language that is not 
esoteric to the field, but easily accessible to someone who commands the 
modern language of mathematics. 


\medskip\noindent
We expect this material to be read online as a key feature are the hyperlinks
giving quick access to internal references spread over many different pages.
If you use an embedded pdf or dvi viewer in your browser, the cross file
links should work.

\medskip\noindent
This project is a collaborative effort and we encourage you to help out.
Please email any typos or errors you find while reading or
any suggestions, additional material, or examples you have to
\href{mailto:automorphic.project@gmail.com}{automorphic.project@gmail.com}.
You can download a tarball containing all source files, extract,
run make, and use a dvi or pdf viewer locally. Please feel free to
edit the LaTeX files and email your improvements.

\medskip\noindent

For convenience, we temporarily list here a description of the parts into which this project will initially be divided, as agreed by the authors:



\section{Attribution}
\label{section-attribution}


\medskip\noindent
Those who contributed to this work are listed on the title page of the book
version of this work and
\href{http://math.jhu.edu/~sakellar/automorphic-project/CONTRIBUTORS}{online}.

\input{chapters}

\bibliography{my}
\bibliographystyle{amsalpha}

\end{document}
